\chapter{Unified Rigidity Reduction}

\section{Abstract Rigidity Template}

\begin{definition}[Rigidity Functional]
Let $(X,\mathcal{R})$ be a refinement system.  
A rigidity functional is a map
\[
\mathcal{F} : X \to \mathbb{R}_{\ge 0}
\]
such that
\begin{enumerate}
\item $\mathcal{F}(x)=0$ characterizes defect-free states,
\item $\mathcal{F}$ is monotone under admissible refinement,
\item $\mathcal{F}$ satisfies a coercive lower bound in terms of entropy or energy.
\end{enumerate}
\end{definition}

\section{Unified Rigidity Lemma}

\begin{theorem}[Unified Rigidity Lemma]
Let $(X,\mathcal{R})$ be an admissible refinement system with bounded capacity $C$ and rigidity functional $\mathcal{F}$.  
If there exists a constant $\alpha>0$ such that
\[
\mathcal{F}(x) \ge \alpha H(x)
\]
for all admissible states, then refinement depth satisfies
\[
\mathrm{ED}(x) \ge \frac{\alpha}{C} H(x).
\]
\end{theorem}

\begin{proof}
Each refinement step reduces entropy by at most $C$.  
Since $\mathcal{F}$ controls entropy from below, eliminating defect requires at least $H(x)/C$ steps up to the multiplicative constant $\alpha$.
\end{proof}

\section{Deterministic Regime}

\begin{theorem}[Deterministic Coercivity Regime]
If $\mathcal{F}$ satisfies a uniform coercive inequality
\[
\mathcal{F}(x) \ge \beta \|x\|^2,
\]
then defect elimination implies a uniform spectral lower bound.
\end{theorem}

\begin{proof}
Coercivity implies stability under perturbation.  
Refinement cannot collapse the defect functional without violating the lower bound.
\end{proof}

\section{Probabilistic Regime}

\begin{theorem}[Averaged Rigidity]
Let $(X_n)$ be random instances with
\[
\mathbb{E}[H(X_n)] = \Theta(n).
\]
If the rigidity functional satisfies
\[
\mathbb{E}[\mathcal{F}(X_n)] \ge \gamma n
\]
for some $\gamma>0$, then
\[
\mathbb{E}[\mathrm{ED}(X_n)] = \Omega(n).
\]
\end{theorem}

\begin{proof}
Apply the entropy–capacity bound in expectation and use linear scaling of entropy.
\end{proof}

\section{Reduction Schema}

\begin{definition}[Reduction]
A problem $\mathcal{P}$ reduces to rigidity if its correctness condition can be expressed as $\mathcal{F}(x)=0$ for a rigidity functional satisfying the Unified Rigidity Lemma.
\end{definition}

\begin{theorem}[Rigidity Reduction Principle]
If $\mathcal{P}$ reduces to rigidity and satisfies bounded-capacity admissibility, then sublinear refinement depth is impossible.
\end{theorem}

\begin{proof}
Immediate from the Unified Rigidity Lemma.
\end{proof}

\section{Discrete and Continuous Bridge}

\begin{theorem}[Discrete–Continuous Correspondence]
Let $(X_n)$ be discrete refinement systems with spectral operators $(A_n)$ in Hilbert spaces.  
If
\[
\langle A_n x, x \rangle \ge \lambda \|x\|^2
\]
uniformly, then the induced rigidity functional satisfies the coercivity condition of the Unified Rigidity Lemma.
\end{theorem}

\begin{proof}
The Rayleigh quotient lower bound induces a uniform defect lower bound, which translates to entropy control via capacity constraints.
\end{proof}

\section{Escape Conditions}

Escaping rigidity requires violation of at least one of:
\begin{itemize}
\item Bounded capacity,
\item Local admissibility (FO$^k$),
\item Coercivity of the rigidity functional,
\item Monotonicity under refinement.
\end{itemize}

\section{Summary}

The Unified Rigidity Reduction identifies a single structural mechanism:

\[
\text{Entropy Growth} + \text{Capacity Bound} + \text{Coercivity}
\;\Longrightarrow\;
\text{Linear Depth}.
\]

All applications in later chapters instantiate this template.
