\chapter{Applications and Reductions}

\section{Purpose of the Applications Chapter}

The preceding chapters define the common URF mechanism:

\[
\text{entropy growth}
+
\text{capacity bound}
+
\text{terminal condition}
\Longrightarrow
\text{refinement depth}.
\]

Applications begin when a domain-specific problem is translated into this
mechanism. A translation is not automatically a proof of the original problem.
It is a reduction surface: it records which objects have been identified, which
bounds have been proved, and which assumptions remain open.

This chapter explains how URF-style reductions appear in computational,
spectral, physical, graph-theoretic, and structural settings.

\section{Reduction Data}

\begin{definition}[Application reduction datum]
An application reduction datum is a tuple
\[
(\mathcal D,X,\mathcal R,H,C,T,\Pi),
\]
where \(\mathcal D\) is the source domain, \(X\) is the translated state space,
\(\mathcal R\) is the admissible refinement class, \(H\) is the entropy or
defect functional, \(C\) is the capacity bound, \(T\) is the terminal set, and
\(\Pi\) is the interpretation map back to the source domain.
\end{definition}

The interpretation map \(\Pi\) is essential. Without it, the URF object may be
internally valid but disconnected from the application it is supposed to
represent.

\begin{definition}[Sound reduction]
A reduction datum is sound if every admissible source-domain process under
study is represented by an admissible refinement trajectory in \(X\), and if
terminal resolution in \(X\) implies the required source-domain conclusion.
\end{definition}

Soundness is the main burden in applications. It is usually harder than the
capacity-depth calculation itself.

\section{Computational Complexity}

In computational applications, the state space \(X\) is often a space of
partial information states: partitions, transcripts, local views, candidate
assignments, proof states, or refinement histories.

\begin{definition}[Computational refinement model]
A computational refinement model consists of a configuration space \(\Omega\),
a refinement state space \(X\), and an admissible update class
\(\mathcal R\) representing the allowed algorithmic steps.
\end{definition}

The entropy \(H\) measures unresolved computational alternatives. For a block
\(B\subseteq \Omega\), a basic entropy is
\[
H(B)=\log |B|.
\]

\begin{theorem}[Computational Depth Transfer]
Suppose a computational problem reduces to a refinement model
\((X,\mathcal R,H,C,T)\). If every admissible update removes at most \(C\)
entropy and every correct solution requires reaching \(T=\{H=0\}\), then every
admissible computation represented by the model has depth at least
\[
\frac{H(x_0)}{C}.
\]
\end{theorem}

\begin{proof}
This is the EntropyDepth lower bound applied to the computational refinement
trajectory.
\end{proof}

\section{SAT-Style Reductions}

For a Boolean search problem on \(n\) variables, the full assignment space has
size \(2^n\). The initial assignment entropy is therefore
\[
\log_2(2^n)=n.
\]

If a local refinement method removes at most \(C\) bits per admissible step,
then terminal isolation of one assignment requires at least \(n/C\) steps.

\begin{remark}[Boundary of SAT-style statements]
This does not prove a SAT lower bound for arbitrary algorithms. It proves a
lower bound only for algorithms already shown to fit the bounded-capacity
refinement model. A theorem about \(P\) versus \(NP\) would require a separate
normalization theorem showing that the relevant algorithms must pass through
such a refinement model.
\end{remark}

\section{Spectral Rigidity}

Spectral applications use operators to measure coercive obstruction.

Let \(L\) be a nonnegative self-adjoint operator with eigenvalues
\[
0=\lambda_0\le \lambda_1\le \lambda_2\le\cdots.
\]

\begin{definition}[Spectral gap]
The first positive spectral gap is a number \(\lambda_1>0\) such that
\[
\langle f,Lf\rangle\ge \lambda_1\|f\|^2
\]
for all \(f\perp \ker L\).
\end{definition}

\begin{theorem}[Spectral Rigidity]
If \(L\) has a positive spectral gap \(\lambda_1>0\), then every
\(f\perp\ker L\) satisfies
\[
\|f\|^2\le \lambda_1^{-1}\langle f,Lf\rangle.
\]
\end{theorem}

\begin{proof}
The spectral gap inequality gives
\[
\langle f,Lf\rangle\ge \lambda_1\|f\|^2.
\]
Dividing by \(\lambda_1>0\) gives the stated estimate.
\end{proof}

This estimate says that deformation away from the kernel has measurable cost.
In URF terms, the spectral gap is a coercivity mechanism.

\section{Physics}

In physical applications, the refinement state may represent accessible
measurements, boundary data, detector outcomes, energy shells, spectral modes,
or coarse-grained macrostates. The capacity bound may come from finite detector
resolution, finite energy, finite channel width, locality, or bounded coupling.

\begin{definition}[Operational physical reduction]
An operational physical reduction represents a physical system by a state
space \(X\), an admissible measurement or update class \(\mathcal R\), an
entropy \(H\), and a terminal condition \(T\) corresponding to the physical
resolution claim.
\end{definition}

\begin{example}[Finite detector capacity]
If a detector has \(B\) distinguishable outcomes per admissible measurement,
then one measurement carries at most \(\log B\) outcome information. If the
unresolved operational entropy is \(H_0\), then at least \(H_0/\log B\)
measurements are required, assuming each measurement is independent in the
required model and terminal resolution requires eliminating that entropy.
\end{example}

A physical application must state the measurement model. Without that model,
the capacity claim is only an analogy.

\section{Graph Structures}

Graph applications often use local neighborhoods, partition refinements, color
refinements, or spectral graph operators.

\begin{definition}[Graph refinement datum]
A graph refinement datum consists of a graph family \(G_n\), a local
observation rule, a partition or label state space, a refinement rule, and a
terminal separation condition.
\end{definition}

If bounded-radius local views leave \(N_n\) globally compatible alternatives,
then the unresolved graph entropy is
\[
H_n=\log N_n.
\]
If each refinement round removes at most \(C\) units of this entropy, then the
terminal depth is at least \(H_n/C\).

\begin{remark}
This is not a graph-isomorphism lower bound unless the refinement model is
proved to capture the graph-isomorphism procedure under study.
\end{remark}

\section{Expanders and Rigidity}

Expanders provide a useful contrast. A graph family with a uniform spectral
gap has strong global connectivity from local incidence data. In URF language,
a gap can act as a coercive control preventing low-cost deformation.

However, expansion alone is not a complete URF lower bound. A full application
must still specify entropy, capacity, terminal resolution, and the admissible
refinement process.

\begin{theorem}[Gap-to-Coercivity Template]
Let \(L_n\) be graph Laplacians with a uniform spectral gap
\[
\lambda_1(L_n)\ge \lambda>0.
\]
Then for every \(f\perp\ker L_n\),
\[
\|f\|^2\le \lambda^{-1}\langle f,L_nf\rangle.
\]
\end{theorem}

\begin{proof}
Apply the spectral gap inequality with \(\lambda_1(L_n)\ge\lambda\).
\end{proof}

\section{Mathematical Structures}

In algebraic, geometric, or categorical settings, an application may use
defect functionals instead of entropy directly. A defect functional
\(F:X\to\mathbb R_{\ge 0}\) becomes URF-relevant when it controls or is
controlled by entropy.

\begin{definition}[Entropy-compatible defect]
A defect functional \(F\) is entropy-compatible if there is a monotone function
\(\psi\) such that
\[
H(x)\le \psi(F(x))
\]
or
\[
F(x)\le \psi(H(x)),
\]
depending on the direction needed by the argument.
\end{definition}

The direction matters. A lower bound on depth requires a quantity that cannot
be eliminated faster than the capacity bound allows.

\section{Reduction Soundness}

The central risk in applications is an unsound translation.

\begin{theorem}[Reduction Soundness Requirement]
A URF lower bound transfers to a source-domain problem only if every
source-domain process in the claimed class is represented by an admissible
refinement trajectory and every source-domain success state maps into the
declared terminal set.
\end{theorem}

\begin{proof}
If a source-domain process is not represented by an admissible trajectory, the
URF lower bound does not apply to it. If source-domain success does not require
entry into the terminal set, the terminal-depth lower bound does not imply a
source-domain lower bound. Therefore both representation and terminal
compatibility are necessary for transfer.
\end{proof}

\section{Application Checklist}

A complete application should provide:

\begin{center}
\begin{tabular}{ll}
\toprule
Field & Required content \\
\midrule
Source domain & problem or system being translated \\
State space & URF representation of source states \\
Admissible process & refinement class modeling source operations \\
Entropy or defect & unresolved quantity being measured \\
Capacity bound & per-step information limit \\
Coercivity & rigidity estimate, if needed \\
Terminal set & translated success condition \\
Soundness bridge & proof that source processes enter the model \\
Boundary & explicit non-claims \\
\bottomrule
\end{tabular}
\end{center}

\section{Boundary}

This chapter gives application templates and reduction requirements. It does
not prove any source-domain lower bound without a soundness bridge, does not
prove SAT hardness, does not prove graph-isomorphism lower bounds, does not
establish any physical theory, and does not prove unrestricted Chronos-RR,
unrestricted H4.1/FGL, P vs NP, or any Clay problem.
