\chapter{EntropyDepth Wall}

\section{Purpose of the EntropyDepth Wall}

The EntropyDepth Wall is the obstruction created when unresolved entropy grows
faster than an admissible refinement process can remove it. It is the combined
form of three earlier ingredients:

\[
\text{initial entropy}
+
\text{bounded step capacity}
+
\text{terminal resolution}
\Longrightarrow
\text{depth lower bound}.
\]

The wall is not a new axiom. It is the point at which entropy accounting
becomes a lower bound on any process inside the declared refinement model.

\section{Refinement Systems}

Let \(X\) be a state space, let \(\mathcal R\) be a class of admissible
refinement maps, and let
\[
H:X\to\mathbb R_{\ge 0}
\]
be an entropy functional.

\begin{definition}[Refinement trajectory]
A refinement trajectory is a finite or infinite sequence
\[
x_0,x_1,x_2,\ldots
\]
such that
\[
x_{t+1}=R_t(x_t)
\]
for some \(R_t\in\mathcal R\).
\end{definition}

\begin{definition}[Terminal entropy state]
A state \(x\in X\) is terminal for entropy \(H\) if
\[
H(x)=0.
\]
The entropy-terminal set is
\[
T_H=\{x\in X:H(x)=0\}.
\]
\end{definition}

\section{Bounded Entropy Loss}

\begin{definition}[Step capacity]
The refinement class \(\mathcal R\) has entropy capacity \(c>0\) if
\[
H(x)-H(Rx)\le c
\]
for every \(x\in X\) and every \(R\in\mathcal R\).
\end{definition}

The capacity \(c\) is a one-step bound. It does not say that states with
entropy larger than \(c\) are inadmissible. It says only that such states
cannot be fully resolved in one admissible step.

\section{Depth Lower Bound}

\begin{theorem}[EntropyDepth Wall]
Let \(x_0\in X\). Suppose every admissible refinement step removes at most
\(c>0\) units of entropy, and suppose terminal resolution requires
\(H(x_t)=0\). Then every admissible terminal trajectory from \(x_0\) has length
at least
\[
\frac{H(x_0)}{c}.
\]
\end{theorem}

\begin{proof}
Let
\[
x_0,x_1,\ldots,x_t
\]
be an admissible trajectory with \(H(x_t)=0\). For each \(i<t\),
\[
H(x_i)-H(x_{i+1})\le c.
\]
Summing these inequalities gives
\[
H(x_0)-H(x_t)\le tc.
\]
Since \(H(x_t)=0\), we obtain
\[
H(x_0)\le tc.
\]
Therefore
\[
t\ge \frac{H(x_0)}{c}.
\]
\end{proof}

\section{Growth Regimes}

The wall becomes meaningful for a family of inputs \(x_n\).

\begin{theorem}[Growth-Regime Transfer]
Let \(g(n)\) be a nonnegative function. Suppose
\[
H(x_n)\ge g(n)
\]
and every admissible refinement step removes at most \(c>0\) entropy, with
\(c\) independent of \(n\). Then
\[
\ED(x_n)\ge \frac{g(n)}{c}.
\]
\end{theorem}

\begin{proof}
Apply the EntropyDepth Wall theorem to \(x_n\) and use
\(H(x_n)\ge g(n)\).
\end{proof}

\begin{center}
\begin{tabular}{lll}
\toprule
Entropy growth & Capacity & Depth lower bound \\
\midrule
\(H(x_n)=\Omega(\log n)\) & \(O(1)\) & \(\Omega(\log n)\) \\
\(H(x_n)=\Omega(n)\) & \(O(1)\) & \(\Omega(n)\) \\
\(H(x_n)=\Omega(n^2)\) & \(O(1)\) & \(\Omega(n^2)\) \\
\(H(x_n)=\Omega(g(n))\) & \(O(1)\) & \(\Omega(g(n))\) \\
\bottomrule
\end{tabular}
\end{center}

Thus the wall is superlinear only when the unresolved entropy is superlinear
relative to the chosen size parameter.

\section{Rigidity Wall}

The EntropyDepth Wall becomes a rigidity wall when unresolved entropy is tied
to a structural defect that cannot be eliminated freely.

\begin{definition}[Rigidity wall datum]
A rigidity wall datum is a tuple
\[
(X,\mathcal R,H,F,c,T),
\]
where \(H\) is entropy, \(F\) is a rigidity or defect functional, \(c\) is a
capacity bound, and \(T\) is the terminal set.
\end{definition}

The functional \(F\) becomes relevant only when it is connected to entropy or
terminal resolution.

\begin{definition}[Entropy--defect bridge]
An entropy--defect bridge is an estimate proving that terminal resolution
requires eliminating a defect measured by \(F\), and that eliminating this
defect requires removing entropy measured by \(H\).
\end{definition}

\begin{theorem}[Rigidity Wall Template]
Assume a rigidity wall datum. Suppose a family \(x_n\) satisfies
\[
H(x_n)\ge g(n).
\]
Suppose terminal resolution requires \(H=0\), and every admissible refinement
step removes at most \(c>0\) entropy. Then
\[
\ED(x_n)\ge \frac{g(n)}{c}.
\]
\end{theorem}

\begin{proof}
The defect data explains why terminal resolution requires removing the entropy
measured by \(H\). Once this terminal requirement is fixed, the EntropyDepth
Wall theorem gives the stated lower bound.
\end{proof}

\section{Wall Versus Bottleneck}

A bottleneck is a local slowdown. A wall is a structural lower bound.

\begin{center}
\begin{tabular}{ll}
\toprule
Object & Meaning \\
\midrule
Bottleneck & a difficult region for a particular process \\
Capacity bound & a one-step information limit \\
Entropy wall & entropy exceeds one-step capacity \\
Rigidity wall & entropy wall tied to structural defect \\
Terminal wall & terminal condition cannot be reached without crossing it \\
\bottomrule
\end{tabular}
\end{center}

The distinction matters because a bottleneck may be bypassed by a different
algorithm, while a wall applies to every process inside the declared admissible
class.

\section{Example: Linear Entropy Wall}

Let
\[
|\Omega_n|=2^n.
\]
Then
\[
H_0=\log_2|\Omega_n|=n.
\]
If each admissible refinement step removes at most \(c\) bits, then
\[
\ED\ge \frac{n}{c}.
\]

This is a linear EntropyDepth Wall.

\section{Example: Quadratic Entropy Wall}

Let
\[
|\Omega_n|=2^{n^2}.
\]
Then
\[
H_0=n^2.
\]
If each admissible refinement step removes at most \(c\) bits, then
\[
\ED\ge \frac{n^2}{c}.
\]

This is a superlinear wall because the initial entropy is superlinear.

\section{Example: Degenerate Capacity}

If the per-step capacity grows with \(n\), the wall weakens. Suppose
\[
H(x_n)=n
\]
but
\[
c_n=n.
\]
Then the bound gives only
\[
\ED(x_n)\ge 1.
\]
Thus a large initial entropy does not force large depth unless capacity is
controlled relative to system size.

\section{Failure Modes}

An EntropyDepth Wall claim fails if:

\begin{enumerate}
\item the entropy lower bound is not proved;
\item the step capacity is not uniform;
\item the terminal condition does not require entropy elimination;
\item the refinement class omits an admissible shortcut;
\item the size parameter is changed after the estimate is stated;
\item the claimed rigidity functional is not connected to entropy.
\end{enumerate}

Each failure mode identifies a missing lemma or assumption.

\section{Audit Checklist}

A complete EntropyDepth Wall proof should provide:

\begin{center}
\begin{tabular}{ll}
\toprule
Field & Required content \\
\midrule
State family & inputs \(x_n\) or configurations \(X_n\) \\
Entropy & explicit functional \(H\) \\
Initial lower bound & estimate \(H(x_n)\ge g(n)\) \\
Refinement class & admissible maps \(\mathcal R\) \\
Capacity & uniform step bound \(c\) \\
Terminal set & condition requiring \(H=0\) or equivalent \\
Depth conclusion & lower bound \(g(n)/c\) \\
Boundary & explicit non-claims \\
\bottomrule
\end{tabular}
\end{center}

\section{Boundary}

This chapter proves only the entropy-capacity form of the EntropyDepth Wall.
It does not prove that any external computational, physical, graph-theoretic,
or spectral process satisfies the required hypotheses unless those hypotheses
are separately supplied. It does not prove unrestricted Chronos-RR,
unrestricted H4.1/FGL, P vs NP, or any Clay problem.
